---
title: "Problem Set 1"
author: "Chase Bookin"
date: "June 24, 2020"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidycensus)
library(ggthemes)
library(ggplot2)
library(janitor)
library(dplyr)
library(gt)
library(readxl)
library(sf)
library(scales)
library(magrittr)
library(haven)
library(infer)
library(Lahman)
library(xml2)
library(rvest)
library(devtools)
library(broom)
library(retrosheet)
```

<h1> **Question 1 ** </h1> 
```{r}

sample_1 <- c(3, 4, 5, 6, 7)
sample_2 <- c(67, 68, 69, 70, 71)

mean_samp1 <- sum(sample_1) / length(sample_1)

sd_samp1 <- round(sqrt(((3-mean_samp1)^2 + (4-mean_samp1)^2 + (5-mean_samp1)^2
                  + (6-mean_samp1)^2 + (7-mean_samp1)^2) / (length(sample_1) - 1)), 3)

mean_samp2 <- sum(sample_2) / length(sample_2)

sd_samp2 <- round(sqrt(((67-mean_samp2)^2 + (68-mean_samp2)^2 + (69-mean_samp2)^2
                  + (70-mean_samp2)^2 + (71-mean_samp2)^2) / (length(sample_2) - 1)), 3)
```

**Sample 1: **  
  - Mean: `r mean_samp1`  
  - Standard Deviation: `r sd_samp1`  

**Sample 2: **   
  -Mean: `r mean_samp2`  
  -Standard Deviation: `r sd_samp2`  
  
The standard deviations of sample 1 and sample 2 are equal. This shows that
although the samples have different means, they are spread around their center
with the same variance.


<h1> **Question 2**  </h1>

```{r}
z_tokyo <- (380000 - 420000) / 20000

z_germany <- (3100 - 3200) / 57

```


Relative to their peers, the worker in Germany is earning more than the worker
in Tokyo. This is demonstrated by the z-score of each worker's salary. The
z-score of the workers' salaries tells the relative position of their salary to
their peers using the mean and standard deviation. In this case, the German
worker's z-score of roughly -1.75 is greater than the Tokyo worker's z-score of
-2, demonstrating that the German worker is earning comparatively more than the
Tokyo worker.


<h1> **Question 3**  

**a) **  
```{r}
z_prob_keane <- 1-.192
z_keane <- 0.87
sd_keane <- (25000 - 21000) / 0.87
```

Standard Deviation: `r round(sd_keane, 2)`

**b) **
```{r}
z_42nd <- -0.2
percentile_42 <- (sd_keane * z_42nd) + 21000
```

42nd Percentile: 20080.46

**c) **
```{r}
# -1.55 <= z <= 1.55
percentile_94 <- (sd_keane * 1.55) + 21000
percentile_06 <- (sd_keane * -1.55) + 21000

```

Middle 88% values: (`r format(round(percentile_06, 2), scientific=FALSE)`, `r format(round(percentile_94, 2), scientific=FALSE)`)






